### 2019-05-23 Datelife_paper1.Rmd
all figures, yay

<!--INTRODUCTION-->
Supersmart is an open tool but not easy to use, still requires a lot of curation and knowledge.

<!--INTRODUCTION: 1st paragraph: general statement that the field knows to be true and important-->
Time of lineage divergence<!--alt lineages: named biological entities, taxon ages?-->
constitutes in many ways the fundamental/main knowledge necessary for evolutionary
understanding.
<!--Along with phylogenetic relationships, it is the most important piece of information
allowing to test alternative evolutionary hypothesis in many areas of research,
from developmental to ecological questions.
Along with migration rates and patterns, it is fundamental to test biogeographical
hypothesis-->
Coupled to species number and distribution, it is the main information necessary
for the study of diversification processes (i.e., the tempo and mode of speciation
and extinction),
<!-- crucial/vital/critical/essential  -->central for the understanding of how biodiversity
patterns are shaped across space, time and clades [@Morlon2014].
Evolutionary understanding also relies on comparative studies, for which knowing
the time context for all life is crucial. Efforts to have a whole tree of life have
been great and here are some examples.
<!--INTRODUCTION: 2nd paragraph: important advances in the field-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods on a very large amount and diversity
of organisms, greatly increasing the quantity of<!--organisms with--> data on taxon
ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of both fossil and molecular-based data on taxon
ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees) names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased
usage by the community<!--total evidence and revbayes (fossilized birth-death)-->,
which coupled to better sharing data practices, are greatly contributing to the
accumulation in number and type of available data on taxon ages.

<!--INTRODUCTION: 3rd paragraph: problems associated-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms
from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
[Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronograms focusing on plants
[PHYLOMATIC], birds [JETZ ET AL BIRDTREE.ORG], and other groups. Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication [@Kumar2017] or the
TimeTree website, though the still substantial chronogram from a previous publication
[@Hedges2015] was made available at OToL.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->

<!--INTRODUCTION: 4th paragraph-->
Despite its great importance, analytical tools to summarize available information
on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of
existing data. First, original chronograms available publicly are scattered across
various repositories (otol tree store, dryad, treebase, journals supplementary data)
usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies
and manual curation of that is usually necessary.
Third, data curation
Recent advances on this area (e.g., supersmart<!--and, which others?-->) aim to:
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of
data curation and knowledge on the group of interest is still necessary. For example,
<!--General issues with dating techniques:--> choosing correct fossils for calibration
requires a lot of expertise and knowledge on the group. An incorrect use of fossils
can generate severe bias in dating results [@Sauquet2012c].
<!--Data curation is necessary at some point (is this always true?)-->
Hence, data curation is still an important part of any biological study. The research
community considers it as an important or even crucial step before data analysis.
Hence, automated processes for large data analysis are frequently received with skepticism.

@@

<!--INTRODUCTION: 5th paragraph: Goal, objectives and predictions-->
Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype `datelife` service over a series
of phylotastic hackathons [@Stoltzfus2013] at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of `datelife`, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
`datelife` is the main service for scaling phylogenetic trees in Phylotastic! system
[@Stoltzfus2013] <!--to construct phylogenetic trees on the fly [http://phylotastic.org/].-->
It can be used through an R package <link to documentation>, a web interface
(<http://www.datelife.org/query/>) and an API<!--(still not up, right?)-->.

>>>>

\newpage
<!--INTRODUCTION: 1st paragraph: general statement that the field knows to be true and important-->
Time of lineage divergence<!--alt lineages: named biological entities, taxon ages?-->
constitutes in many ways a fundamental piece of information for evolutionary
understanding.
<!--Along with phylogenetic relationships, it is crucial for testing alternative
evolutionary hypothesis in many areas of research,
from developmental to ecological questions @Felsenstein1985a @Webb2008.
It has been shown that even a little bit of information on a lineages time of origin is good to study ecological stuff @Webb2008
comparative studies, in general, for which knowing
the time context for all life is crucial.
Along with migration rates and patterns, it is fundamental to test biogeographical
hypothesis-->
Coupled to species number and distribution, it allows studying the tempo and mode of speciation
and extinction),
<!-- crucial/vital/critical/essential --> central for understanding of how biodiversity
patterns are shaped across space, time and clades [@Morlon2014].
<!-- Efforts to have a whole tree of life have
been great and here are some examples. -->
<!--INTRODUCTION: 2nd paragraph: important advances in the field-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods on a very large amount and diversity
of organisms, greatly increasing the quantity and quality of<!--organisms with--> data on taxon
ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of
<!-- both fossil and molecular-based  -->
data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
<!-- pbdb ??-->
OToL alone hosts now more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees) names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased
usage by the community<!--total evidence and revbayes (fossilized birth-death)-->,
which coupled to better sharing data practices, are greatly contributing to the
accumulation in number and type of available data on taxon ages.

<!--INTRODUCTION: 3rd paragraph: problems associated-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms
from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
[Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronogram repositories focusing on plants
[@webb2005phylomatic], birds [@Jetz2012], and fish . Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication [@Kumar2017] or the
TimeTree website, though the still substantial chronogram from a previous publication
[@Hedges2015] was made available at OToL.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->

<!--INTRODUCTION: 4th paragraph-->
Despite its great importance, analytical tools to summarize available information
on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of
existing data. First, original chronograms available publicly are scattered across repositories
<!-- (otol tree store, dryad, treebase,  -->
and journals supplementary data, usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies
and manual curation of that is usually necessary.
Third, data curation
Recent advances on this area (e.g., supersmart<!--and, which others?-->) aim to:
Supersmart is an open tool but not easy to use, still requires a lot of curation and knowledge.
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of
data curation and knowledge on the group of interest is still necessary. For example,
<!--General issues with dating techniques:--> choosing correct fossils for calibration
requires a lot of expertise and knowledge on the group. An incorrect use of fossils
can generate severe bias in dating results [@Sauquet2012c].
<!--Data curation is necessary at some point (is this always true?)-->
Hence, data curation is still an important part of any biological study. The research
community considers it as an important or even crucial step before data analysis.
Hence, automated processes for large data analysis are frequently received with skepticism.

@@

<!--INTRODUCTION: 5th paragraph: Goal, objectives and predictions-->
Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype `datelife` service over a series
of phylotastic hackathons [@Stoltzfus2013] at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of `datelife`, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
`datelife` is the main service for scaling phylogenetic trees in Phylotastic! system
[@Stoltzfus2013] <!--to construct phylogenetic trees on the fly [http://phylotastic.org/].-->
It can be used through its R package <link to documentation>, and a web interface
(<http://www.datelife.org/query/>)<!-- and an API (still not up, tho)-->.

### 2019-05-23 Datelife_paper1.tex
all figures, yay

Supersmart is an open tool but not easy to use, still requires a lot of curation and knowledge.

Time of lineage divergence
constitutes in many ways the fundamental/main knowledge necessary for evolutionary
understanding.
Coupled to species number and distribution, it is the main information necessary
for the study of diversification processes (i.e., the tempo and mode of speciation
and extinction),
central for the understanding of how biodiversity
patterns are shaped across space, time and clades (Morlon \protect\hyperlink{ref-Morlon2014}{2014}).
Evolutionary understanding also relies on comparative studies, for which knowing
the time context for all life is crucial. Efforts to have a whole tree of life have
been great and here are some examples.
In the past two decades, the possibility to obtain good quality DNA sequences
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods on a very large amount and diversity
of organisms, greatly increasing the quantity of data on taxon
ages across the tree of life.
To date, there is a large amount of both fossil and molecular-based data on taxon
ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms.
Methods to include living and fossil lineages are in continued development and increased
usage by the community,
which coupled to better sharing data practices, are greatly contributing to the
accumulation in number and type of available data on taxon ages.

The TimeTree project (Hedges et al. \protect\hyperlink{ref-Hedges2006}{2006}, \protect\hyperlink{ref-Hedges2015}{2015}; Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}) has aggregated chronograms
from 3,163 studies, encompassing 97,085 species (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}), and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236\% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
{[}Mayr2010; Moran2011{]}). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronograms focusing on plants
{[}PHYLOMATIC{]}, birds {[}JETZ ET AL BIRDTREE.ORG{]}, and other groups. Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}) or the
TimeTree website, though the still substantial chronogram from a previous publication
(Hedges et al. \protect\hyperlink{ref-Hedges2015}{2015}) was made available at OToL.

Despite its great importance, analytical tools to summarize available information
on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of
existing data. First, original chronograms available publicly are scattered across
various repositories (otol tree store, dryad, treebase, journals supplementary data)
usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies
and manual curation of that is usually necessary.
Third, data curation
Recent advances on this area (e.g., supersmart) aim to:
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of
data curation and knowledge on the group of interest is still necessary. For example,
choosing correct fossils for calibration
requires a lot of expertise and knowledge on the group. An incorrect use of fossils
can generate severe bias in dating results (Sauquet et al. \protect\hyperlink{ref-Sauquet2012c}{2012}).
Hence, data curation is still an important part of any biological study. The research
community considers it as an important or even crucial step before data analysis.
Hence, automated processes for large data analysis are frequently received with skepticism.

@@

Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype \texttt{datelife} service over a series
of phylotastic hackathons (Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}) at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of \texttt{datelife}, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
\texttt{datelife} is the main service for scaling phylogenetic trees in Phylotastic! system
(Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013})
It can be used through an R package , a web interface
(\url{http://www.datelife.org/query/}) and an API.

@@

Data from source chronograms was also used to date tree topologies with no branch length information and trees with branch lengths in relative substitution rates (Figs. \ref{fig:cvbladj} and \ref{fig:cvbold}). As a form of cross validation, we used tree topologies from each study and calibrated them using information from all other source chronograms. In the absence of branch length data, the ages of internal nodes were approximately recovered in almost all cases (except for studies 3, and 5; Fig. \ref{fig:cvbladj}). Maximum tree ages were only approximately recovered in one case (study 2; Fig. \ref{fig:cvbladj}).
Branch lengths were successfully generated using the BOLD database for all source chronograms. However, dating with PATHd8 (using congruified calibrations) was only successful in\\
three cases (studies 3, 5, and 9; Fig. \ref{fig:cvbold}). From these, two trees have a different sampling than the original source chronogram, mainly because DNA data for some species is absent from the BOLD. Maximum ages are quite different from source chronograms, but this might be explained also by the differences in sampling between source chronograms and BOLD trees.
More examples and details can be consulted in \url{https://github.com/LunaSare/datelife_examples}.

@@

@@

\begin{center}
\textsc{Figure \ref{fig:cvbold}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees with branch length reconstructed with data from the Barcode of Life Database (BOLD) dated using PATHd8. We could construct a tree with branch lengths for all source chronograms. However, dating with PATHd8 was only successfull in three source chronograms shown here.

>>>>

Time of lineage divergence
constitutes in many ways a fundamental piece of information for evolutionary
understanding.
Coupled to species number and distribution, it allows studying the tempo and mode of speciation
and extinction),
central for understanding of how biodiversity
patterns are shaped across space, time and clades (Morlon \protect\hyperlink{ref-Morlon2014}{2014}).
In the past two decades, the possibility to obtain good quality DNA sequences
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods on a very large amount and diversity
of organisms, greatly increasing the quantity and quality of data on taxon
ages across the tree of life.
To date, there is a large amount of
data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
OToL alone hosts now more than 200 chronograms.
Methods to include living and fossil lineages are in continued development and increased
usage by the community,
which coupled to better sharing data practices, are greatly contributing to the
accumulation in number and type of available data on taxon ages.

The TimeTree project (Hedges et al. \protect\hyperlink{ref-Hedges2006}{2006}, \protect\hyperlink{ref-Hedges2015}{2015}; Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}) has aggregated chronograms
from 3,163 studies, encompassing 97,085 species (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}), and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236\% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
{[}Mayr2010; Moran2011{]}). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronogram repositories focusing on plants
(Webb and Donoghue \protect\hyperlink{ref-webb2005phylomatic}{2005}), birds (Jetz et al. \protect\hyperlink{ref-Jetz2012}{2012}), and fish . Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}) or the
TimeTree website, though the still substantial chronogram from a previous publication
(Hedges et al. \protect\hyperlink{ref-Hedges2015}{2015}) was made available at OToL.

Despite its great importance, analytical tools to summarize available information
on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of
existing data. First, original chronograms available publicly are scattered across repositories
and journals supplementary data, usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies
and manual curation of that is usually necessary.
Third, data curation
Recent advances on this area (e.g., supersmart) aim to:
Supersmart is an open tool but not easy to use, still requires a lot of curation and knowledge.
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of
data curation and knowledge on the group of interest is still necessary. For example,
choosing correct fossils for calibration
requires a lot of expertise and knowledge on the group. An incorrect use of fossils
can generate severe bias in dating results (Sauquet et al. \protect\hyperlink{ref-Sauquet2012c}{2012}).
Hence, data curation is still an important part of any biological study. The research
community considers it as an important or even crucial step before data analysis.
Hence, automated processes for large data analysis are frequently received with skepticism.

@@

Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype \texttt{datelife} service over a series
of phylotastic hackathons (Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}) at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of \texttt{datelife}, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
\texttt{datelife} is the main service for scaling phylogenetic trees in Phylotastic! system
(Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013})
It can be used through its R package , and a web interface
(\url{http://www.datelife.org/query/}).

@@

Data from source chronograms was also used to date tree topologies with no branch length information and trees with branch lengths in relative substitution rates (Figs. \ref{fig:cvbladj} and \ref{fig:cvbold}). As a form of cross validation, we used tree topologies from each study and calibrated them using information from all other source chronograms. In the absence of branch length data, the ages of internal nodes were approximately recovered in almost all cases (except for studies 3, and 5; Fig. \ref{fig:cvbladj}). Maximum tree ages were only approximately recovered in one case (study 2; Fig. \ref{fig:cvbladj}).
Branch lengths were successfully generated using the BOLD database for all source chronograms. However, dating with PATHd8 (using congruified calibrations) was only successful in
three cases (studies 3, 5, and 9; Fig. \ref{fig:cvbold}). From these, two trees have a different sampling than the original source chronogram, mainly because DNA data for some species is absent from the BOLD. Maximum ages are quite different from source chronograms, but this might be explained also by the differences in sampling between source chronograms and BOLD trees.
More examples and details can be consulted in \url{https://github.com/LunaSare/datelife_examples}.

@@

\leavevmode\hypertarget{ref-webb2005phylomatic}{}%
Webb C.O., Donoghue M.J. 2005. Phylomatic: Tree assembly for applied phylogenetics. Molecular Ecology Notes. 5:181--183.

@@

\begin{center}
\textsc{Figure \ref{fig:cvbold}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees with branch length reconstructed with data from the Barcode of Life Database (BOLD) dated using PATHd8. We could construct a tree with branch lengths for all source chronograms. However, dating with PATHd8 was only successful in three source chronograms shown here.
